---
title: "MA329_PROJ"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  some questions to lm???

1. use lm() of full model to see the p-value (or use plot() to see if each factor have some multidimensional connection?)(what about some have intersection?) & stepwise/other methods to find the reduced one(then test whether the reduced model is better)

2. high leverage pts(delete or not)
test normality/ independence/ linearity/ homoscedasticity. (parallel or progressive relationship)

3. the difference between ANOVA & summary

4. data visualization (without model, see the outline of data?)

5. corrective methods: delecting observations, transforming variables, adding or deleting variables, using another regression approach...?


#  import packages
```{r, warning = FALSE}
library(readxl)
library(MASS)
library(car)
library(base)
```


#  inport data
```{r}
data <- read_excel("data.xlsx")
```

#  missing points detection
```{r}
na_data = is.na(data)
table(na_data)
```
no missing pts.

#  boxplot of all factors(maybe we don't need to do this)
```{r}
par(mfrow=c(2,5))
boxplot(data$Aids, main = "Aids")
boxplot(data$agriculture, main = "Agriculture")
boxplot(data$fertility,main = "Fertility")
boxplot(data$GovSpend ,main = "GovSprend")
boxplot(data$infantMortality ,main = "infantMortality")
boxplot(data$sanitation ,main = "sanitation")
boxplot(data$smoke_F ,main = "smoke_F")
boxplot(data$smoke_M ,main = "smoke_M")
boxplot(data$urbanPop ,main = "urbanPop")
```
We can see some outliers?

#  summary of all factor, to see their min max and mean etc.
```{r}
summary(data)

```
what can we find?

#  see the full model and find the significant one.
```{r}
attach(data)
fullmodel <- lm(Aids ~ (agriculture + fertility + GovSpend + infantMortality + sanitation + smoke_F + smoke_M + urbanPop ))
summary(fullmodel)
anova(fullmodel)
aov(fullmodel)
```
as p-value is very small, so we reject the null H0... so the model is meaningful, aids and other factors have connection.

#  see the scatterplotmatrix but it's too big
```{r}
data_num <- data[c("agriculture", "fertility", "GovSpend", "infantMortality", "sanitation", "smoke_F", "smoke_M", "urbanPop")]
car::scatterplotMatrix(data_num)
```

it's too much that maybe we don't need to see this test?

#  full model's test
```{r}
round(cor(data_num),4)
full_matrix <- model.matrix(fullmodel)[,-1]
e <- eigen(t(full_matrix) %*% full_matrix)
e$val
sqrt(e$val[1] / e$val)
#round(vif(full_matrix),3)
library(ggcorrplot)
cor_heart <- cor(data[,2:10])
cor_heart
ggcorrplot(cor_heart,lab = F)
plot(fit2)
```



#  use stepwise to determine the significant factors
```{r}
stepwisemethod <- as.data.frame(data[, c("Aids", "agriculture", "fertility", "GovSpend", "infantMortality", "sanitation", "smoke_F", "smoke_M", "urbanPop")])
fit <- lm(Aids ~ (agriculture + fertility + GovSpend + infantMortality + sanitation + smoke_F + smoke_M + urbanPop ))
stepAIC(fit, direction = "both")
```
then we can get this: lm(formula = Aids ~ agriculture + infantMortality + urbanPop), also fit the full model's significancity.


#  test VIF
```{r}
reducedmodel <- lm(log(Aids) ~ agriculture + fertility + urbanPop)
summary(reducedmodel)
library(faraway)
red_matrix <- model.matrix(reducedmodel)[,-1]
vif(red_matrix)
```
all are not larger than 10.


#  test normality
```{r}
qqPlot(reducedmodel, labels=row.names(data$Country), id.method="identify",simulate=TRUE, main="Q-Q Plot")
#states["80",]
#fitted(fit)["80"]
#residuals(fit)["80"]
par(mfrow = c(1,2))
qqPlot(fit, labels=row.names(data$Country), id.method="identify",simulate=TRUE, main="Q-Q Plot of full model")
logfit <- lm(log(Aids) ~ (agriculture + fertility + GovSpend + infantMortality + sanitation + smoke_F + smoke_M + urbanPop ))
qqPlot(logfit, labels=row.names(data$Country), id.method="identify",simulate=TRUE, main="Q-Q Plot of LOG transformation")
#states["80",]
#fitted(fit)["80"]
#residuals(fit)["80"]
summary(logfit)
```

#  test independence of errors
```{r}
durbinWatsonTest(fit)
```

#  test linearity
```{r}
crPlots(fit)
```

#  test homoscedasticity
```{r}
ncvTest(fit)
spreadLevelPlot(fit)
```

#  outlier
```{r}
outlierTest(fit)
```


QQPlot
```{r}
par(mfrow = c(1,2))
logreducedmodel <- lm(log(Aids) ~ agriculture + infantMortality + urbanPop)
qqPlot(fullmodel, labels=row.names(data$Country), id.method="identify",simulate=TRUE, main="Q-Q Plot of model")
qqPlot(logfit, labels=row.names(data$Country), id.method="identify",simulate=TRUE, main="Q-Q Plot of reduced model with LOG transformation")
```

```{r}
ncvTest(fit2)
spreadLevelPlot(logreducedmodel)
durbinWatsonTest(fit2)
outlierTest(fit2)
summary(fit2)
```

```{r}
library(car)
attach(data)
par(mfrow = c(3,3))
scatter.smooth(rstudent(fit2) ~ agriculture, col = "grey", las = 1, ylab = "Studendized residuals")
scatter.smooth(rstudent(fit2) ~ fertility, col = "grey", las = 1, ylab = "Studendized residuals")
scatter.smooth(rstudent(fit2) ~ urbanPop, col = "grey", las = 1, ylab = "Studendized residuals")
```

```{r}
summary(powerTransform(data$Aids))
```


```{r}
summary(logreducedmodel)
```










```{r}
stepwisemethod <- as.data.frame(data[, c("Aids", "agriculture", "fertility", "GovSpend", "smoke_F", "smoke_M", "urbanPop")])
fit <- lm(log(Aids) ~ (agriculture + fertility + GovSpend + smoke_F + smoke_M + urbanPop ))
stepAIC(fit, direction = "both")
fit2 <-lm(formula = log(Aids) ~ agriculture + fertility + urbanPop)
summary(fit2)
```

```{r}
library(flextable)
hist(data$Aids)
A <- data.frame(data[6,])
a <- rbind(A,data[73,])
a

data[order(data$Aids, decreasing = TRUE),]
```

```{r}
data_low <- read_excel("data_low.xlsx")
data_high <- read_excel("data_high.xlsx")
par(mfrow = c(1,2))
boxplot(data_low$Aids, main = "low GovSpend countries' aid's rate")
boxplot(data_high$Aids, main = "high GovSpend countries' aid's rate")
```

```{r}
attach(data_high)
stepwisemethod <- as.data.frame(data_low[, c("Aids", "agriculture", "fertility", "GovSpend", "infantMortality", "sanitation", "smoke_F", "smoke_M", "urbanPop")])
fit <- lm(log(Aids) ~ (agriculture + fertility + GovSpend + infantMortality + sanitation + smoke_F + smoke_M + urbanPop ))
stepAIC(fit, direction = "both")

```



```{r}
data_new <- read_excel("data_new.xlsx")
library(lmPerm)
library(multcomp)
attach(data_new)
exang <- as.factor(GovSpendDifference)
fit=aovp(Aids~GovSpendDifference, perm="Prob", iter = 500000000)
summary(fit)
```